---
title: "Quarto Cheatsheet"
format: html
---

Based on materials from [RstatsTextbook](https://elenlefoll.github.io/RstatsTextbook/13_LiterateProgramming.html).


::: {.callout-note}
This guide is designed for complete beginners to Quarto and R. Each example includes detailed explanations of what the code does and why you'd use it.
:::

# Quarto Cheatsheet

## Basic Setup

### Loading Libraries
## 🏁 Getting Started

### Loading a Package
```{r}
#| label: setup
#| include: false
library(checkdown)
```

**What this does:**
- `library(checkdown)` loads the checkdown package, which adds interactive quiz and exercise features to your document.
- `#| label: setup` gives this chunk a name (like a bookmark) so you can reference it later.
- `#| include: false` tells Quarto: "Run this code, but don't show the code or any output in the final document."

**When to use this:**
- At the beginning of your document to load any packages you'll need.
- For setup code that doesn't need to be visible to readers.

## Data Import and Basic Operations

### Loading Data
### Importing Data from CSV Files
```{r}
#| label: import-data
#| include: false
L1.data <- read.csv("data/L1_data.csv")
L2.data <- read.csv("data/L2_data.csv")
```

**What this does:**
- `read.csv()` reads data from a CSV file into R.
- `L1.data` and `L2.data` are names we're giving to our data (like labeled boxes to store things in).
- The file paths in quotes tell R where to find the data files.

**Important notes:**
- The file path is relative to your Quarto document's location.
- Make sure your data files are in a `data` folder in the same folder as your Quarto document.
- You can also use full file paths like `C:/Users/YourName/Documents/data.csv` (Windows) or `/Users/YourName/Documents/data.csv` (Mac).

### Basic Data Exploration
### Basic Data Exploration
```{r}
# Basic summary statistics
summary(L1.data$Age)

# Count observations by group
table(L1.data$Gender)
```

**Understanding the code:**
1. `summary(L1.data$Age)`
   - `L1.data$Age` means "get the Age column from the L1.data dataset"
   - `summary()` calculates basic statistics: minimum, 1st quartile, median, mean, 3rd quartile, and maximum
   - Example output:
     ```
     Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
     17.0    25.0    35.0    38.0    50.0    65.0
     ```

2. `table(L1.data$Gender)`
   - Counts how many of each category exist in the Gender column
   - Example output:
     ```
       F   M   
      48  42
     ```
     (This means there are 48 females and 42 males)

**When to use:**
- Use `summary()` to quickly understand the distribution of numerical data.
- Use `table()` to count categories in categorical data (like gender, age groups, etc.).

## Data Visualization

### Basic Plot
## 📊 Creating Basic Plots

### Making a Scatter Plot
```{r}
#| label: fig-scatterplot
#| fig-cap: "L2 participants' lexical proficiency in English and their professional occupational group"
#| fig-height: 5
#| fig-asp: 0.618
#| message: false
#| echo: fenced

plot(L2.data$LexTALE, L2.data$GrammarR,
     xlab = "Lexical Proficiency",
     ylab = "Grammar Score",
     pch = 16, col = "steelblue")
```

**Understanding the code:**
- `plot(x, y)` creates a basic scatter plot where:
  - `x` is the horizontal axis (LexTALE scores)
  - `y` is the vertical axis (GrammarR scores)
- `xlab` and `ylab` add labels to the axes
- `pch = 16` makes solid circle points (try numbers 1-25 for different shapes)
- `col = "steelblue"` sets the point color (you can use color names or hex codes like "#4682B4")

**Chunk options explained:**
- `label: fig-scatterplot` - Names this figure for cross-referencing
- `fig-cap` - Adds a caption below the figure
- `fig-height: 5` - Makes the plot 5 inches tall
- `fig-asp: 0.618` - Sets the golden ratio for width-to-height
- `message: false` - Hides any messages from R
- `echo: fenced` - Shows the code in a simple format without syntax highlighting

### Interactive Plot with plotly
## 🔄 Creating Interactive Plots

### Interactive Plot with plotly
```{r}
#| label: fig-scatterplot-plotly
#| fig-cap: "An interactive plot of L2 participants' lexical proficiency in English"
#| message: false
#| warning: false
#| code-fold: true
#| code-summary: "Show R code to generate the interactive plot below."

library(plotly)
plot_ly(data = L2.data, 
        x = ~LexTALE, 
        y = ~GrammarR,
        type = "scatter",
        mode = "markers",
        text = ~paste("Age: ", Age, "<br>Group: ", Group))
```

**What makes this special:**
- Viewers can hover over points to see details
- Zoom in/out of specific areas
- Download the plot as an image
- Pan around the plot

**Code breakdown:**
- `library(plotly)` - Loads the plotly package (install it first with `install.packages("plotly")` if needed)
- `plot_ly()` - Creates an interactive plot
  - `data = L2.data` - Uses the L2.data dataset
  - `x` and `y` specify which columns to plot
  - `type = "scatter"` makes a scatter plot
  - `mode = "markers"` shows individual points
  - `text` creates hover information (shows Age and Group when you hover over a point)

**Chunk options explained:**
- `code-fold: true` - Hides the code by default (readers can click to show it)
- `code-summary` - Customizes the "Show code" button text

**Try this:**
1. Hover over points to see the Age and Group
2. Click and drag to zoom into an area
3. Double-click to reset the view

## Tables

### Basic Table with kable
## 📋 Creating Tables

### Making a Simple Table
```{r}
#| label: tbl-participants
#| tbl-cap: "Participant Demographics"
#| echo: fenced

knitr::kable(
  data.frame(
    Group = c("L1", "L2"),
    N = c(nrow(L1.data), nrow(L2.data)),
    Mean_Age = c(mean(L1.data$Age), mean(L2.data$Age))
  ),
  digits = 1
)
```

**What this does:**
Creates a table that looks like this:

| Group |  N | Mean_Age |
|-------|----|----------|
| L1    | 90 |     38.2 |
| L2    | 67 |     35.8 |

**Code breakdown:**
1. `data.frame()` creates a table with three columns:
   - `Group`: Contains "L1" and "L2"
   - `N`: Counts rows in each dataset with `nrow()`
   - `Mean_Age`: Calculates average age for each group

2. `knitr::kable()` formats the table nicely
   - `digits = 1` rounds numbers to 1 decimal place

**Chunk options explained:**
- `label: tbl-participants` - Names the table for referencing
- `tbl-cap` - Adds a caption above the table
- `echo: fenced` - Shows the code in a simple format

**Tip:** Add `format = "html"` inside `kable()` for fancier HTML tables with features like sorting and searching.

## Inline Code

### Basic Inline Code
## ✏️ Using Inline Code

### Inserting Values into Text
```markdown
The mean age of L1 participants is `r mean(L1.data$Age) |> round(1)` years.
```

**What this does:**
- When rendered, it will show something like: "The mean age of L1 participants is 38.2 years."
- The actual number is calculated automatically from your data

**How it works:**
- Everything between `r` and `}` is R code that gets run
- `mean(L1.data$Age)` calculates the average age
- `|> round(1)` rounds the result to 1 decimal place (the pipe `|>` means "then do this"). This is the new base R pipe operator (introduced in R 4.1.0).

**Why this is great:**
- If your data changes, the numbers in your text update automatically
- No more copy-paste errors in your report!

---

## References

This guide was inspired by and adapted from:
- [Literate Programming with R Markdown](https://elenlefoll.github.io/RstatsTextbook/13_LiterateProgramming.html)

**Try this:**
1. Change the `1` in `round(1)` to `0` or `2` and see what happens
2. Try calculating and displaying other statistics in your text

## Cross-Referencing

### Referencing Figures/Tables
## 🔗 Cross-Referencing

### Referencing Figures and Tables
```markdown
As shown in Figure @fig-scatterplot, there is a positive correlation...

See Table @tbl-participants for demographic information.
```

**What this does:**
- Creates clickable links to your figures and tables
- Automatically numbers everything for you
- If you add/remove items, the numbers update automatically

**How to use it:**
1. Make sure your figure/table has a label (like `fig-scatterplot` or `tbl-participants`)
2. Reference it in your text using `@label`
   - For figures: `@fig-label`
   - For tables: `@tbl-label`

**Example output in your document:**
"As shown in Figure 1, there is a positive correlation..."

**Pro tip:** Use the Visual Editor in RStudio to insert these references using the @ menu - it's much easier!

## Conditional Content

### Showing/Hiding Content
## 🎚️ Conditional Content

### Showing/Hiding Content Based on Conditions
```{r}
#| echo: false
#| eval: 1 == 1  # Change to FALSE to hide this content

cat("This content will only appear if the condition is TRUE")
```

**What this does:**
- The code inside will only run if the condition after `eval:` is TRUE
- In this case, `1 == 1` is always TRUE, so the content will show
- Change it to `FALSE` to hide the content

**Practical examples:**
1. Show different content in draft vs. final versions:
   ```
   #| eval: params$draft == TRUE
   cat("DRAFT VERSION - Not for distribution")
   ```

2. Include extra details in an appendix:
   ```
   #| eval: params$include_appendix == TRUE
   cat("Additional technical details...")
   ```

**How to use parameters:**
Add this to your document's YAML header (at the very top):
```yaml
---
params:
  draft: true
  include_appendix: false
---
```

## Advanced Features

### Parameterized Reports
## 🎛️ Creating Dynamic Reports with Parameters

### Setting Up Parameters
Add this to the very top of your document (in the YAML header):
```yaml
---
title: "My Report"
params:
  show_code: true    # Toggle code visibility
  dataset: "L1"      # Choose which dataset to use
  highlight: "blue"  # Set highlight color
---
```

### Using Parameters in Your Code
```{r}
#| echo: params$show_code

# Use different data based on parameter
if(params$dataset == "L1") {
  data <- L1.data
  title <- "Native Speakers (L1)"
} else {
  data <- L2.data
  title <- "Second Language Learners (L2)"
}

# Create a plot with the selected data
plot(data$Age, main = title, col = params$highlight)
```

**How to use this:**
1. Click "Render" with different parameter values to create different versions
2. Or click the arrow next to "Render" and select "Render with Parameters" to choose values

**Real-world uses:**
- Create one report that works for different departments
- Generate student-specific reports
- Make quick comparisons between different datasets

## 🏆 Best Practices for Beginners

### 1. Organizing Your Document
```markdown
# Data Preparation
(Chunks for importing and cleaning data)

# Analysis
(Chunks for statistical tests and calculations)

# Visualizations
(Chunks for creating plots and tables)
```
- **Why?** Makes your document easier to navigate and understand
- **Tip:** Use Markdown headers to create clear sections

### 2. Making Your Work Reproducible
```{r}
# At the beginning of your document
set.seed(123)  # Makes random results consistent
sessionInfo()  # Records package versions
```
- **Why?** So others (or future you) can get the same results
- **Tip:** Include a "Session Info" section at the end

### 3. Keeping Things Fast
```{r}
#| cache: true
# Long-running calculation here
result <- long_running_function()
```
- **Why?** Saves time by not re-running slow calculations
- **Tip:** Give cached chunks meaningful names

### 4. Making Your Code Readable
```{r}
# Good chunk label
#| label: clean-survey-data

# Bad chunk label
#| label: chunk1
```
- **Why?** Helps you find things later
- **Tip:** Use verbs in your chunk names (e.g., "import-data", "clean-names", "plot-results")

### 5. Getting Help
- Type `?function_name` in the R console for help
- Use the "Help" tab in RStudio
- Search online with "R" and what you're trying to do

Remember: Everyone was a beginner once! Don't be afraid to make mistakes - they're how we learn. Happy coding! 🚀
