---
title: "Quarto Chunk Options Reference"
format: html
---

Based on materials from [RstatsTextbook](https://elenlefoll.github.io/RstatsTextbook/13_LiterateProgramming.html).

# Quarto Chunk Options Reference

## Code Display Options

### Code Execution
- `eval: true/false`  
  Controls whether the code in the chunk is executed. When `false`, the code is shown but not run.

- `echo: true/false/"fenced"`  
  Controls code visibility:  
  - `true`: Shows code with syntax highlighting (default)  
  - `false`: Hides the code completely  
  - `"fenced"`: Shows code in plain text without syntax highlighting, useful for documentation

- `include: true/false`  
  Determines if the chunk's output is included in the final document. When `false`, the code still runs but output is suppressed.

### Message Control
- `message: true/false`  
  Controls display of messages (e.g., package loading messages)

- `warning: true/false`  
  Controls display of warnings

- `error: true/false`  
  Controls whether to show errors or stop execution

## Figure Options

### Basic Figure Settings
- `fig-cap: "caption text"`  
  Adds a caption to the figure

- `fig-cap-location: top/bottom`  
  Positions the figure caption at the top or bottom

- `fig-height: number`  
  Sets the figure height in inches (e.g., `5` for 5 inches)

- `fig-width: number`  
  Sets the figure width in inches

- `fig-asp: number`  
  Sets the aspect ratio (height/width) of the figure (e.g., `0.618` for golden ratio)

- `fig-align: "center"`  
  Aligns the figure (center, left, or right)

## Table Options

- `tbl-cap: "caption text"`  
  Adds a caption to the table

- `tbl-cap-location: top/bottom`  
  Positions the table caption at the top or bottom

- `tbl-colwidths: [width1, width2, ...]`  
  Specifies relative column widths (e.g., `[80,20]` for 80% and 20%)

## Code Folding

- `code-fold: true/false`  
  Enables/disables the ability to show/hide code

- `code-summary: "text"`  
  Custom text to display for the "Show code" button when code is folded

## Chunk Identification

- `label: "chunk-name"`  
  Names the chunk for cross-referencing (e.g., `fig-` prefix for figures, `tbl-` for tables)

## Common Use Case Examples

### Basic Setup Chunk
```{r}
#| label: setup
#| include: false
#| message: false
#| warning: false
# Your setup code here
```

### Figure with Caption
```{r}
#| label: fig-myplot
#| fig-cap: "My figure caption"
#| fig-height: 5
#| fig-asp: 0.618
# Your plotting code here
```

### Table with Caption
```{r}
#| label: tbl-mytable
#| tbl-cap: "My table caption"
#| tbl-cap-location: top
#| tbl-colwidths: [70,30]
# Your table code here
```

### Interactive Plot with Code Folding
```{r}
#| label: fig-interactive
#| fig-cap: "Interactive plot"
#| message: false
#| warning: false
#| code-fold: true
#| code-summary: "Show R code to generate the interactive plot below."
# Your interactive plotting code here
```

## Best Practices

1. Always use meaningful labels for chunks you want to reference
2. Use `include: false` for setup chunks that don't need to show output
3. Be consistent with figure and table numbering (use `fig-` and `tbl-` prefixes)
4. Use code folding for long code sections to keep the document clean
5. Set appropriate figure dimensions to ensure good readability

For more detailed information, refer to the [Quarto documentation](https://quarto.org/).

